Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States

### 2024-11-04 paper/paper.qmd
Added model summary.

>>>>

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Subset the data for each candidate
harris_data <- analysis_data %>%
  filter(candidate_name == "Kamala Harris")

trump_data <- analysis_data %>%
  filter(candidate_name == "Donald Trump")

# Create model summary data for Harris
modelsumharris <- harris_data %>%
  mutate(
    fit = predict(model3, newdata = ., interval = "confidence"),
    lower = fit[, "lwr"],
    upper = fit[, "upr"],
    fitted = fit[, "fit"]
  ) %>%
  select(end_date, fitted, lower, upper, state)  # Include state column

# Create model summary data for Trump
modelsumtrump <- trump_data %>%
  mutate(
    fit = predict(model6, newdata = ., interval = "confidence"),
    lower = fit[, "lwr"],
    upper = fit[, "upr"],
    fitted = fit[, "fit"]
  ) %>%
  select(end_date, fitted, lower, upper, state)  # Include state column

# Combine the datasets for plotting
modelsumharris$candidate_name <- "Kamala Harris"
modelsumtrump$candidate_name <- "Donald Trump"
combined_data <- bind_rows(modelsumharris, modelsumtrump)
```
```{r}

#| echo: false
#| warning: false
#| message: false
#| label: fig-modsum1
#| eval: true
#| fig-cap: 'Model summaries for Kamala Harris and Donald Trump at the national level using polling data.'

ggplot(combined_data, aes(x = as.Date(end_date), y = fitted, color = candidate_name, fill = candidate_name)) +
  geom_smooth(method = "loess", se = TRUE, span = 0.2) +
  scale_color_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
  scale_fill_manual(values = c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred")) +
  labs(x = "Date", y = "Support (%)") +
  theme_minimal() +
  theme(legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  facet_wrap(~ candidate_name, scales = "free_y")

```
```{r}

#| echo: false
#| warning: false
#| message: false
#| label: fig-modsum2
#| eval: true 
#| fig-cap: 'Model summary of Donald Trump’s support trends across key battleground states.'
# Define the target states
target_states <- c("Arizona", "Georgia", "Michigan", "Pennsylvania", "Wisconsin", "North Carolina", "Nevada")

# Filter combined_data for Donald Trump, exclude national data, and limit to target states
trump_state_data <- combined_data %>%
  filter(candidate_name == "Donald Trump", state %in% target_states)

# Plotting
ggplot(trump_state_data, aes(x = as.Date(end_date), y = fitted, fill = candidate_name)) +
  geom_smooth(method = "loess", se = TRUE, span = 0.2, color = "indianred") +
  scale_fill_manual(values = c("Donald Trump" = "indianred")) +
  labs(x = "Date", y = "Support (%)",
       title = "Support trends for Donald Trump across different states") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  facet_wrap(~ state, scales = "free_y")  # Separate subplots for each state

```
```{r}

#| echo: false
#| warning: false
#| message: false
#| label: fig-modsum1
#| fig-cap: 'Model summary of Kamala Harris’s support trends across key battleground states.'
# Define the target states
target_states <- c("Arizona", "Georgia", "Michigan", "Pennsylvania", "Wisconsin", "North Carolina", "Nevada")

# Filter combined_data for Kamala Harris and limit to target states
harris_state_data <- combined_data %>%
  filter(candidate_name == "Kamala Harris", state %in% target_states)

# Plotting
ggplot(harris_state_data, aes(x = as.Date(end_date), y = fitted, fill = candidate_name)) +
  geom_smooth(method = "loess", se = TRUE, span = 0.2, color = "steelblue") +
  scale_fill_manual(values = c("Kamala Harris" = "steelblue")) +
  labs(x = "Date", y = "Support (%)",
       title = "Support trends for Kamala Harris across different states") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 0),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  ) +
  facet_wrap(~ state, scales = "free_y")  # Separate subplots for each state
